---
title: "social-media-dataset-simulation"
output: html_document
date: "2023-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(dagitty)
```


## Introduction

In this document, we generate the dataset used as an example for the multiverse analysis. 
Note: I do not recall the seed used for the original analysis, hence the dataset, and the results of the multiverse analysis may vary slightly.

The dataset we are going to simulate will have the relationships outlined in the DAG below:

```{r}
g <- dagitty('dag {
    I1 [pos="0.5,0"]
    I2 [pos="1.5,0"]
    I3 [pos="2.5,0"]
    I4 [pos="3.5,0"]
    parent_income [pos="0,2"]
    sm_usage [pos="2,1"]
    depression [pos="6,1"]
    activity [pos="4,2"]
    has_sibling [pos="4,3"]
    age [pos="4,4"]
    
    sm_usage -> I1
    sm_usage -> I2
    sm_usage -> I3
    sm_usage -> I4
    parent_income -> sm_usage
    sm_usage -> depression
    sm_usage -> activity
    activity -> depression
    activity -> has_sibling
    sm_usage -> has_sibling
    depression -> has_sibling
    age -> sm_usage
    age -> depression
    age -> has_sibling
}')
plot(g)
```

These variables are:

- I1: self reported number of hours spent on social media per week
- I2: self reported likert scale rating of social media usage (1 [very low/no usage at all] - 5 [very high usage])
- I3: self reported number of friends/followers on social media across all platforms
- I4: self reported number of posts shared in a week
- self-reported age
- self reported number of hours per week engaged in physical activity or exercise of any kind
- Binary choice indicating whether participant has siblings
- Age of siblings (if any)
- Total parental income (rounded to the nearest $10k)

## Simulation

First, we define the correlation between all of the variables which impact one another:

```{r}
N = 300

b_income_sm = 0.5
b_sm_depression = 0.4
b_sm_I1 = 0.7
b_sm_I2 = 0.7
b_sm_I3 = 0.15
b_sm_I4 = 0.15
b_sm_activity = 0.2
b_depression_activity = -0.7
b_sm_activity = -0.3
b_age_sm = -0.15
b_age_depression = -0.2
b_siblings_sm = 0.2
b_siblings_depression = -0.1
b_siblings_activity = 0.3
```

We can now simulate the complete dataset. We adopt a two step approach when simulating. First, we define every independent variable as a standard normal distribution (i.e. mean = 0; sd = 1). We then calculate the dependent variables.

```{r}
set.seed(123)
social_media_scaled = tibble(
  parent_income = rnorm(N),
  age = rnorm(N),
  has_siblings = rbinom(N, 1, 0.72),
  social_media = parent_income*b_income_sm + 
                  age*b_age_sm +
                  has_siblings * b_siblings_sm +
                  rnorm(N, 0, sd = sqrt(1 - (b_income_sm^2 + b_age_sm^2 + b_siblings_sm^2))),
  I1 = social_media*b_sm_I1 + rnorm(N, 0, sd = sqrt(1 - b_sm_I1^2)),
  I2 = social_media*b_sm_I2 + rnorm(N, 0, sd = sqrt(1 - b_sm_I2^2)),
  I3 = social_media*b_sm_I3 + rnorm(N, 0, sd = sqrt(1 - b_sm_I3^2)),
  I4 = social_media*b_sm_I4 + rnorm(N, 0, sd = sqrt(1 - b_sm_I4^2)),
  depression = social_media*b_sm_depression +
              age*b_age_depression +
              has_siblings * b_siblings_depression +
              rnorm(N, 0, sd = sqrt(1 - (b_sm_depression^2 + b_age_depression^2 + b_siblings_depression^2))),
  physical_activity = social_media*b_sm_activity +
              has_siblings * b_siblings_activity +
              depression * b_depression_activity +
              rnorm(N, 0, sd = sqrt(1 - (b_sm_activity^2 + b_siblings_activity^2 + b_depression_activity^2)))
)
```

Finally, we scale the variables appropriately:

```{r}
social_media = social_media_scaled |> 
  # we transform all the variables to the appropriate scale
  mutate(
    parent_income = round(exp(parent_income*0.2 + 4.5)) * 1000,
    age = round(rescale(age, to = c(11, 17))),
    I1 = exp(I1*0.3 + 3),
    I2 = round(rescale(I2, to = c(1, 7))),
    I3 = round(exp(I3*0.5 + 6)),
    physical_activity = round(exp(rescale(physical_activity, c(0, 2.5))))
  )

head(social_media)
```




